---
title: "R Notebook"
output: word_document
editor_options: 
  chunk_output_type: console
---

#To Update
```{r}
project_name="Project_NP0084-MB4_5_pt6_2"
manifest_name="metadata_Summary_1.csv"
otu_name="otu_Summary_1.csv"
sample_depth = 20000
```

#Load required libraries
############### 
```{r}
library("biom")
library(data.table)
library(ggplot2)
library(tm)
library(vegan)
library("biom")
library(tidyr)
library(ggplot2)
library(cowplot)
library(scales)
library(phyloseq)
library(qiime2R)
library(tibble)
library(gridExtra)
library(tidyverse)
```

#Analysis
###############
#1. Taxonomy Reference - Expected frequencies
```{r}
#Read in controls taxonomy file, grouped by Genus
control_expect <- read.table("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Controls\\Taxonomy\\Taxonomy_Genus_2019.txt",sep="\t",header=TRUE)
rownames(control_expect) <- unique(control_expect[,"Genus"])
```

#2. Load OTU and metadata tables
```{r}
#Create location link
project_location = paste("T:\\DCEG\\Projects\\Microbiome\\CGR_MB\\MicroBiome\\",project_name,sep="")

#Read in database with variable information
metadata_baseline <- read.table(paste(project_location,"\\Output\\R\\OTUs\\",manifest_name,sep=""),sep=",",header = TRUE)
rownames(metadata_baseline)<-metadata_baseline$X
metadata_baseline<-metadata_baseline[-1]

otu_baseline_complete <- read.table(paste(project_location,"\\Output\\R\\OTUs\\",otu_name,sep=""),sep=",",header = TRUE)
rownames(otu_baseline_complete)<-otu_baseline_complete$X
otu_baseline_complete<-otu_baseline_complete[-1]

rownames(metadata_baseline)==rownames(otu_baseline_complete) #Should be true
```

#3. OTU Counts by sample
Read in OTU information, and determine total OTUs
```{r}
genus_found<-colnames(otu_baseline_complete)
samplelist_seqcontrols<-rownames(otu_baseline_complete)

#For each sample, determine the total read counts
for (a in samplelist_seqcontrols){
 #Set sum to 0
 sum=0
 
 #For each genus
 for (b in genus_found){
  
  #Add the number of counts together
  sum <- otu_baseline_complete[a,b] + sum
 }
 
 #When complete, push the sum to a new "Total_OTUs" row
 otu_baseline_complete[a,"Total_OTUs"] <- sum
}
```

#4. Observed to Expected Counts - Seq Controls
```{r}
samplelist_baseline_seqcontrol <- subset(metadata_baseline_filtered, (metadata_baseline_filtered$Sample.Type == "SeqControl"))

subjectlist_baseline_seqcontrol <- subset(metadata_baseline_filtered, (metadata_baseline_filtered$Sample.Type == "SeqControl"))
subjectlist_baseline_seqcontrol <- unique(subjectlist_baseline_seqcontrol$SubjectID)

observed_count =0

#Individual Samples
for (a in subjectlist_baseline_seqcontrol){
 genus_temp <- subset(control_expect, !(control_expect[,a] == "A"))
 genus_temp <- genus_temp$Genus
  
 for (b in samplelist_baseline_seqcontrol){
  subject.id <- metadata_baseline_filtered[b,"SubjectID"]
  
  if(a==subject.id){
   
   for (c in genus_temp){
    otu_count <- otu_baseline_filtered[b,c]
    
    if(!(is.na(observed_count) | is.null(otu_count) | otu_count ==0)){
     observed_count = observed_count+1
    }
    
   }
   
  }
  
  observed_baseline_seqcontrols[b,"Observed"] <- observed_count
  observed_baseline_seqcontrols[b,"Expected"] <- length(genus_temp)
  observed_baseline_seqcontrol[b,"SubjectID"] <- subject.id
  observed_count =0
 
 }
 
}

write.csv(observed_baseline_seqcontrols,"/datatables/observed_baseline_seqcontrols.csv")

#Average by SubjectID
count =0
for (a in subjectlist_baseline_seqcontrol){
 for (b in samplelist_baseline_seqcontrol){
  subject.id <- observed_baseline_seqcontrols[b,"SubjectID"]
  
  if(a=subject.id){
   observed_count<-observed_baseline_seqcontrols[b,"Observed"]+observed_count
   expected_count<-observed_baseline_seqcontrols[b,"Expected"]
   count = count+1
  }
 }
 observed_baseline_seqcontrols[a,"Observed"]<-observed_count/count
 observed_baseline_seqcontrols[a,"Expected"]<-expected_count
 observed_baseline_seqcontrol[a,"SubjectID"]<-a
 count =0
}

write.csv(observed_baseline_seqcontrols,"/datatables/observed_baseline_seqcontrols.csv")

```

#8. Relative Abundance  - Seq Controls
```{r}
#Complete Relative Abundance
for (a in subjectlist_baseline_seqcontrol){
 
 for (b in ncol(otu_baseline_filtered)-1){
   relab <- otu_baseline_filtered[a,b]
   
   if(!(is.na(relab) | is.null(relab))){
     relab_baseline_seqcontrols[a,b] <- otu_baseline_filtered[a,b]/otu_baseline_filtered[a,"Total_OTUs"]
   }

 }
 
}
write.csv(relab_baseline_seqcontrols,"/datatables/relabun_baseline_seqcontrols.csv")

#Expected relative abundance
for (a in subjectlist_baseline_seqcontrol){
 genus_temp <- subset(control_expect, !(control_expect[,a] == "A"))
 genus_temp <- genus_temp$Genus
  
 for (b in samplelist_baseline_seqcontrol){
  subject.id <- metadata_baseline_filtered[b,"SubjectID"]
  
  if(a==subject.id){
   
   for (c in genus_temp){
    rel_abund_observed <- relab_baseline_seqcontrols[b,c]
    rel_abund_expected <- control_expect[c,a]
    gram_stain <- control_expect[c,"Gram"]
    
    relab_baseline_seqcontrols_expected[b,paste(c,".Obs",sep="")]<-rel_abund_observed
    relab_baseline_seqcontrols_expected[b,paste(c,".Exp",sep="")]<-rel_abund_expected
    relab_baseline_seqcontrols_expected[b,"GramStain"]<-gram_stain
    relab_baseline_seqcontrols_expected[b,"SubjectID"]<-a
   }
   
  }

 }
 
}

write.csv(relab_baseline_seqcontrols_expected,"/datatables/relab_baseline_seqcontrols_expected.csv")

#Expected Average by SubjectID

count=0
observed_count=0

for (a in subjectlist_baseline_seqcontrol){
 genus_temp <- subset(control_expect, !(control_expect[,a] == "A"))
 genus_temp <- genus_temp$Genus
 
 for (b in genus_temp){
  
  for (c in samplelist_baseline_seqcontrol){
   
   subject.id <- relab_baseline_seqcontrols_expected[b,"SubjectID"]
   
   if(a=subject.id){
    observed_count<-relab_baseline_seqcontrols_expected[c,paste(b,".Obs",sep="")]+observed_count
    expected_count<-relab_baseline_seqcontrols_expected[c,paste(b,".Exp",sep="")]
    gram_stain <- relab_baseline_seqcontrols_expected[c,"GramStain"]
    count = count+1
    
   }
   
  }
  
  relab_baseline_seqcontrols_expected[a,paste(b,".Obs",sep="")]<-observed_count/count
  relab_baseline_seqcontrols_expected[a,paste(b,".Exp",sep="")]<-expected_count
  relab_baseline_seqcontrols_expected[a,"GramStain"]<-gram_stain
  relab_baseline_seqcontrols_expected[a,"SubjectID"]<-a
  count=0
  observed_count=0
 }
}
  
write.csv(relab_baseline_seqcontrols_expected,"/datatables/relab_baseline_seqcontrols_expected.csv")

#Unexpected relative abundance
for (a in subjectlist_baseline_seqcontrol){
 genus_temp <- subset(control_expect, !(control_expect[,a] == "A"))
 genus_temp <- genus_temp$Genus
 
 genus_found <- relab_baseline_seqcontrols$Genus
 genus_temp <- subset(genus_found,!genus_temp$V1)
  
 for (b in samplelist_baseline_seqcontrol){
  subject.id <- metadata_baseline_filtered[b,"SubjectID"]
  
  if(a==subject.id){
   
   for (c in genus_temp){
    rel_abund_observed <- relab_baseline_seqcontrols[b,c]
    gram_stain <- control_expect[c,"Gram"]
    
    relab_baseline_seqcontrols_unexpected[b,paste(c,".Obs",sep="")]<-rel_abund_observed
    relab_baseline_seqcontrols_unexpected[b,"GramStain"]<-gram_stain
    relab_baseline_seqcontrols_unexpected[b,"SubjectID"]<-a
   }
   
  }

 }
 
}

write.csv(relab_baseline_seqcontrols_unexpected,"/datatables/relab_baseline_seqcontrols_unexpected.csv")

#Expected Average by SubjectID

count=0
observed_count=0

for (a in subjectlist_baseline_seqcontrol){
 genus_temp <- subset(control_expect, !(control_expect[,a] == "A"))
 genus_temp <- genus_temp$Genus
 
 genus_found <- relab_baseline_seqcontrols$Genus
 genus_temp <- subset(genus_found,!genus_temp$V1)
 
 for (b in genus_temp){
  
  for (c in samplelist_baseline_seqcontrol){
   
   subject.id <- relab_baseline_seqcontrols_unexpected[b,"SubjectID"]
   
   if(a=subject.id){
    observed_count<-relab_baseline_seqcontrols_unexpected[c,paste(b,".Obs",sep="")]+observed_count
    gram_stain <- relab_baseline_seqcontrols_unexpected[c,"GramStain"]
    count = count+1
    
   }
   
  }
  
  relab_baseline_seqcontrols_unexpected[a,paste(b,".Obs",sep="")]<-observed_count/count
  relab_baseline_seqcontrols_unexpected[a,"GramStain"]<-gram_stain
  relab_baseline_seqcontrols_unexpected[a,"SubjectID"]<-a
  count=0
  observed_count=0
 }
}
  
write.csv(relab_baseline_seqcontrols_unexpected,"/datatables/relab_baseline_seqcontrols_unexpected.csv")

```

#9 Relative Abundance % Difference
```{r}
for (a in subjectlist_baseline_seqcontrol){
 
 genus_temp <- subset(control_expect, !(control_expect[,a] == "A"))
 genus_temp <- genus_temp$Genus
  
 for (b in genus_temp){
  observed_count<-relab_baseline_seqcontrols_expected[a,paste(b,".Obs",sep="")]
  expected_count<-relab_baseline_seqcontrols_expected[a,paste(b,".Exp",sep="")]
  
   if(!(is.na(observed_count)|is.null(observed_count)|is.na(observed_count)|is.null(observed_count)|observed_count=0|expected_count=0)){

    relab_baseline_seqcontrols_percdiff[a,b]<- ((observed_count-expected_count)/expected_count)*100
   }
 }
}

write.csv(relab_baseline_seqcontrols_percdiff,"/datatables/relab_baseline_seqcontrols_percdiff.csv")

```

###############to update###############
#11 Plot graph - Actual to Expected - IntraControl Variability
For each kit, show variability, and map
```{r}
#Create list for plots, start counter
plot_list <- list()
i=1

for (a in control_types){
   #Create genus list of specfic control
   genus_temp <- control_expect[!is.na(control_expect[,a]),]
   genus_temp <- (genus_temp[,"Genus"])
   
   #Create a list of samples that are of the specified control type
   sample_list<- subset(metadata_baseline_filtered, Type==a)
   sample_list <- unique(sample_list$SampleID)

   #Create dataframe
   temp<-data.frame()
   count=1
   
   #For each Genus in list, determine the samples value
   for (b in genus_temp){
    
     #For each of the samples in the sample list
     for(c in sample_list){
      temp[count,"Genus"]<-b
      temp[count,"Obs_Per"]<- samples_obs_abund[b,c]
      temp[count,"Sample"]<-c
      temp[count,"Exp_Per"]<-control_expect[b,a]
      count=count+1
     }
   }
  
    graph_title <- paste (a,": Intracontrol Variablity",sep = "")

    #Plot Data
    graph_new <- ggplot(data=temp,mapping=aes(x=Obs_Per,y=Exp_Per,color=Sample))+
     geom_point()+ #Can add points to graph
     geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
     labs(title=graph_title, x="Observed Rel. Abundance", y= "Expected Rel. Abundance")+
     geom_text(mapping=aes(x=Obs_Per,y=Exp_Per,label=Genus),check_overlap = TRUE)
     
    
    plot_list[[i]] = graph_new
    i=i+1
}

#Save all plots
for (i in 1:length(control_types)){
 
 #ID which control is being run
 a <- control_types[i]
 
 #Save the file
 file_name = paste("Data/Graphs/Intracontrol Variability_",a, ".tiff", sep="")
 tiff(file_name)
 print(plot_list[[i]])
 dev.off()
}
```

#12 Plot graph - Actual to Expected - One graph
For each kit, show variability, and map
```{r}
#Create list for plots, start counter
temp<-data.frame()
count<-1
for (a in control_types){
   #Create genus list of specfic control
   genus_temp <- control_expect[!is.na(control_expect[,a]),]
   genus_temp <- (genus_temp[,"Genus"])
   
   #Create a list of samples that are of the specified control type
   sample_list<- subset(metadata_baseline_filtered, Type==a)
   sample_list <- unique(sample_list$SampleID)

   #For each Genus in list, determine the samples value
   for (b in genus_temp){
    
     #For each of the samples in the sample list
     for(c in sample_list){
      temp[count,"Genus"]<-b
      temp[count,"Obs_Per"]<- samples_obs_abund[b,c]
      temp[count,"Sample"]<-c
      temp[count,"Exp_Per"]<-control_expect[b,a]
      temp[count,"Type"] <- a
      count=count+1
     }
   }
}

#Plot graph
graph_new<- ggplot(data=temp,mapping=aes(x=Obs_Per,y=Exp_Per,color=Type))+
 #geom_point()+ #Can add points to graph
 geom_smooth(method = "lm",se=FALSE)+ #Add a smooth line
 labs(title="Sequencing Control Comparison", x="Observed Rel. Abundance", y= "Expected Rel. Abundance")
graph_new
write.csv(temp,"temp2.csv")

file_name =("Data/Graphs/Kit Differences_ZE.tiff")
tiff(file_name)
print(graph_new)
dev.off()


```