---
title: "9. DB Comparisons_Seq"
author: "Sevilla"
date: "October 12, 2019"
output: word_document
editor_options: 
  chunk_output_type: console
---
############### 
#Project Overview
To determine impact of updated Silva Database has on sequening control performance, using TAR and TDR as metrics.

#QIIME2 Pipeline
Version 1 of the pipeline was used (T:\DCEG\Projects\Microbiome\CGR_MB\MicroBiome\sc_scripts_qiime2_pipeline\working) for this test. Three runs were performed using the sample dataset on three different start days for a stochastic test. All runs were the same, and so R_Run1 was chosen for this analysis. Data was analyzed using Silva119, GreenGenes, and an internally trained Silva132. Output data from the QIIME2 pipeline is stored on the T Drive (T:\DCEG\Projects\Microbiome\CGR_MB\MicroBiome\Project_NP0084_MB4)

#Sample Set
Samples included in this anaysis were the sequencing controls of NP0084-MB4, MB5, and MB6, merged into one run. This includes a 43 controls samples of seven different mock communities and two blank-types. Blank-types were subsequently removed.
-Project-ID: NP0084-MB4, NP0084-MB5, NP0084-MB6
-Sample-Cat: MSA, Zymo.Seq
-Sample-Descrip: MSA1000, MSA1001, MSA1002, MSA1003, D6305, D6306, D6311

#Analysis Overview
Phyloseq objects will be adapted from the analysis pipeline created for the fecal analysis of NP0084-MB4,MB5,MB6. This includes 1. Initial Processing which created the PhyloSeq objects from the three flowcells, and merged them together. OTU tables at two taxonomic levels (family and genus) were created during this step. and 2. Baseline_Seq which imported these taxonomic tables and merged the expected control information to created observed/expected data tables. Output data is stored on the T Drive (T:\DCEG\CGF\TechTransfer\Microbiome\Extraction\Optimization\Fecal\Fresh Fecal Optimization_2017.08\Phase I\Analysis\NP0084-MB4,5,6\R_Seq).

OTU tables will be compared below 

#Required libraries
```{r}
library("biom")
library(ape)
library(data.table)
library(ggplot2)
library(tm)
library(vegan)
library(tidyr)
library(cowplot)
library(scales)
library(phyloseq)
library(qiime2R)
library(tibble)
library(gridExtra)
source("sources/miseqR.R")
library(tidyverse)
source("sources/ggrare.R") #github library: https://rdrr.io/github/gauravsk/ranacapa/
library(compareDF)
library(VennDiagram)
```

#Input
```{r}
#Adapted from 1. Initital Processing code for fecal analysis
data_dir = c("T:\\DCEG\\Projects\\Microbiome\\CGR_MB\\MicroBiome\\")
project_name=c("Project_NP0084_MB4\\")
run_list=c("DB_Variations")
manifest_name=c("NP0084-MB4_08_29_19_metadata_seq.txt")

output_location_sto=c("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\NP0084-MB4,5,6\\R_Stochastic\\")
output_location_seq=c("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Analysis\\NP0084-MB4,5,6\\R_DB_Seq\\")

sample_depth=100000
reference_db=c("silva","silva132","greengenes")

#Adapted from 2. Baseline_Seq
control_location=c("T:\\DCEG\\CGF\\TechTransfer\\Microbiome\\Extraction\\Optimization\\Fecal\\Fresh Fecal Optimization_2017.08\\Phase I\\Controls\\Taxonomy\\")

#New Code Requirements
output_run_list=c("R_Run1")
taxa_levels=c("Family","Genus")
```


############### Adapted from 1. Initital Processing code for fecal analysis
#Directory Creation
```{r}
#Create directories - One for each output location, and the following subdirectories:
sub_create<-c("OTUs","Taxa","Graphs","Summary")

for (a in sub_create){
  dir.create(paste(output_location_seq,a,sep=""))
}

```

#Create PhySeq Objects
```{r}
count=1
for (a in reference_db){
  run_data_dir = paste(data_dir,project_name,run_list,sep="")
 
  #Read OTUS
  otus<-read_qza(paste(run_data_dir,"\\Output\\qza_results\\table_dada2_qza_merged_parts_final\\table_dada2_merged_final_filt.qza",sep=""))
  
  #Read rooted tree
  tree<-read_qza(paste(run_data_dir,"\\Output\\qza_results\\phylogeny_qza_results\\rooted_tree.qza",sep=""))
  
  #Read taxonomy file
  taxonomy<-read_qza(paste(run_data_dir,"\\Output\\qza_results\\taxonomy_qza_results\\taxonomy_",a,".qza",sep=""))
  
  #Edit table
  tax_table<-do.call(rbind, strsplit(as.character(taxonomy$data$Taxon), ";"))
  
  col_list<-c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
  if(!(a=="greengenes")){
   for (i in 8:ncol(tax_table)){
    col_list[[i]]<-paste("Subspecies",i,sep="") #number of columns varies depending on database
   }
  }
  
  colnames(tax_table)<-col_list
  rownames(tax_table)<-taxonomy$data$Feature.ID
  
  #read metadata
  metadata<-read.table(paste(run_data_dir,manifest_name,sep="\\"),sep='\t', header=T, row.names=1, comment="")
  
  #Create phylo object
  phyloname<-paste("physeq_complete_",a,sep="")
  assign(phyloname,phyloseq(otu_table(otus$data, taxa_are_rows = T), phy_tree(tree$data), tax_table(tax_table), sample_data(metadata)))
  
  phylo_number=count #sets the number of phyloseq objects that will need to be merged downstream
  count=count+1
}
```

#Prune taxonmoy
```{r}
#Prune for bacteria only
for (a in reference_db){
 phy_name<-paste("physeq_complete_",a,sep="")
 phy_filt_name<-paste("physeq_filt_",a,sep="")
  
 if(!(a=="greengenes")){
  
  assign(phy_filt_name,get(phy_name) %>%
   subset_taxa(
     Kingdom == "D_0__Bacteria" &
     Family  != "D_0__Bacteria; D_1__Proteobacteria; D_2__Alphaproteobacteria; D_3__Rickettsiales; D_4__mitochondria" &
     Class   != "D_0__Bacteria; D_1__Cyanobacteria; D_2__Chloroplast"
   ))
 } else{
   assign(phy_filt_name,get(phy_name) %>%
   subset_taxa(
     Kingdom == "k__Bacteria" &
     Family  != "k__Bacteria; p__Proteobacteria; c__Alphaproteobacteria; o__Rickettsiales; f__mitochondria" &
     Class   != "k__Bacteria; p__Cyanobacteria; c__Chloroplast"
   ))
  
 }

 #Print summaries
 s<-summary(sample_data(get(phy_filt_name))$Sample.Cat)
 capture.output(s, file = paste(output_location_seq,"Summary\\summary_prefilter_sample_",a,".Cat.txt",sep=""))
 s<-summary(sample_data(get(phy_filt_name))$Sample.Descrip)
 capture.output(s, file = paste(output_location_seq,"Summary\\summary_prefilter_sample_",a,".Descrip.txt",sep=""))
 
}

```

#Filter taxa >.001, filter samples with less than 10000 read
```{r}
#NOTE: Samples should have been filtered during Q2 pipeline - done to ensure this was completed
for (a in reference_db){
 phy_filt_name<-paste("physeq_filt_",a,sep="")
 
 assign(phy_filt_name, filter_taxa(get(phy_filt_name), function(x) mean(x) > 1e-2, TRUE))
 assign(phy_filt_name, prune_samples(sample_sums(get(phy_filt_name)) > 10000, get(phy_filt_name)))
 assign(phy_filt_name, get(phy_filt_name) %>% scale_reads(n=sample_depth))

 
 o<-otu_table(get(phy_name))
 write.csv(o, file = paste(output_location_seq,"OTUs\\otu_table_filt_",a,".csv",sep=""))
}

```

#Merge OTU's into taxa tables
```{r}
for (a in reference_db){
 phy_filt_name<-paste("physeq_filt_",a,sep="")

 for (b in taxa_levels){
  
  # Create a factor corresponding to the taxalevel
  taxfac = factor(tax_table(get(phy_filt_name))[, b])
  
  # Tabulate the counts for each genera in each sample
  taxtab = apply(otu_table(get(phy_filt_name)), MARGIN = 2, function(x) {
   tapply(x, INDEX = taxfac, FUN = sum, na.rm = TRUE, simplify = TRUE)
  })
  taxtab<-as.data.frame(taxtab)
  
  for (c in rownames(taxtab)){
   taxa<-c
   #Greengenes and silva have different taxonomic naming schemes - must address each
   #If silva
   if(!a=="greengenes"){
    
    if(grepl("Unassigned",taxa)){ #Silva has unassigned tax 
     taxtab[c,b]<-"Unassigned"
      
     #Genus and family levels will be named different (genus: D5, family: D4)
     #If Genus
    } else if(b=="Genus"){
     if(!grepl("D_5__",taxa)){ #if the name includes a D_ and number less than 5, it will be from a higher tax level
      taxtab[c,b]<-"HigherGenus"
     } else{
      colname_update<-gsub("D_\\d__","",taxa) #Need to remove the formatting of taxonmy for each viewing downstream
      taxtab[c,b]<-colname_update
     }
     #If not Genus
    } else
     if(!grepl("D_4__",taxa)){
      taxtab[c,b]<-"HigherFamily"
     } else{
      colname_update<-gsub("D_\\d__","",taxa)
      taxtab[c,b]<-colname_update
     }
    #If greengenes
    } else{    
     if(b=="Genus"){
      if(!grepl("g__",taxa)){
       taxtab[c,b]<-"HigherGenus"
      } else{
       colname_update<-str_remove(c, "g__") #Need to remove the formatting of taxonmy for each viewing downstream
       colname_update<-gsub("[","",colname_update,fixed=TRUE) #fixed = TRUE disables regex
       colname_update<-gsub("]","",colname_update,fixed=TRUE)
       taxtab[c,b]<-colname_update
      }
     #If higher than genus
     } else{
      if(!grepl("f__",taxa)){
       taxtab[c,b]<-"HigherFamily"
      } else{
       colname_update<-str_remove(c, "f__") #Need to remove the formatting of taxonmy for each viewing downstream
       colname_update<-gsub("[","",colname_update,fixed=TRUE) #fixed = TRUE disables regex
       colname_update<-gsub("]","",colname_update,fixed=TRUE)
       taxtab[c,b]<-colname_update
      }
     }
    }
  }
  
  taxtab<-aggregate(taxtab[-ncol(taxtab)],by=list(taxtab[,b]),FUN="sum") #use -ncol since first col should not be summed
   
  #Check if there is a blank first column because any unassigned taxa (originally was "g__") will be blank due to above
  if(taxtab[1,1]==" "){
   taxtab[1,1]<-"Unknown"
  }
  
  taxtab<-t(taxtab) #transpose for downstream metadata matching
  colnames(taxtab)<-taxtab[1,]
  taxtab<-taxtab[-1,]
  colnames(taxtab) <- gsub(" ", "", colnames(taxtab)) #Remove any spaces in names
   
  file_name =paste(output_location_seq,"Taxa\\taxa_",b,"_",a,".csv",sep="")
  write.csv(taxtab,file_name)
   
  metatab <- as.data.frame(sample_data(get(phy_filt_name)))
  file_name =paste(output_location_seq,"Taxa\\metadata_Summary_",b,"_",a,".csv",sep="")
  write.csv(metatab,file_name)
 }
}
```

################## New Code
#Read in tax tables
```{r}
for (a in reference_db){
 for (b in taxa_levels){
  tax_table_name<-paste("taxa",b,a,sep="_")
  assign(tax_table_name,read.csv(paste(output_location_seq,"Taxa\\taxa_",b,"_",a,".csv",sep=""),row.names = 1,check.names = FALSE))
 }
}
```

#Compare tax tables
```{r}
#Determine which taxa are in all db's, in silva db's only, and in silva119 or silva132
for (a in taxa_levels){
 #lengths of each
 a1=length(colnames(get(paste("taxa",a,"silva",sep="_"))))
 a2=length(colnames(get(paste("taxa",a,"silva132",sep="_"))))
 a3=length(colnames(get(paste("taxa",a,"greengenes",sep="_"))))
 
 #intersection list
 i12<-intersect(colnames(get(paste("taxa",a,"silva",sep="_"))),colnames(get(paste("taxa",a,"silva132",sep="_"))))
 i23<-intersect(colnames(get(paste("taxa",a,"silva132",sep="_"))),colnames(get(paste("taxa",a,"greengenes",sep="_"))))
 i13<-intersect(colnames(get(paste("taxa",a,"silva",sep="_"))),colnames(get(paste("taxa",a,"greengenes",sep="_"))))
 i123<-intersect(i12,colnames(get(paste("taxa",a,"greengenes",sep="_"))))
 
 #intersection values
 n12=length(i12)
 n23=length(i23)
 n13=length(i13)
 n123=length(i123)
 
 #Draw and save venn diagram
 ven_plot<-draw.triple.venn(a1,a2,a3,n12,n23,n13,n123,category = c("Silva119","Silva132","GreenGenes"),fil=c("blue","red","green"))
 final<-grid.arrange(gTree(children=ven_plot),top=paste("Venn Diagram of Tax at ",a," level",sep=""))
 
 tiff(paste(output_location_seq,"\\Graphs\\venn_",a,".tiff",sep=""),compression="lzw")
 grid.draw(final)
 dev.off()
}

```

#OTU differences between databases
```{r}
taxa_diffs<-function(input1,input2){
 temp<-data.frame()
 rowcount=1
 for (a in rownames(input1)){
  for (b in colnames(input1)){
   temp[rowcount,"Source.Descrip"]<-metadata[a,"Source.Descrip"]
   temp[rowcount,"PercDiff"]<-((input1[a,b]-input2[a,b])/input1[a,b])*100
   temp[rowcount,"Taxa"]<-b
   temp[rowcount,"SampleID"]<-a
   rowcount=rowcount+1
  }
 }
 return(temp)
}

bar_percdiff<-function(df1,genus,dbs,controltype){
 if(controltype=="MSA"){
  controls<-c("MSA1000","MSA1001","MSA1002","MSA1003")
 } else{
  controls<-c("D6305","D6306","D6311")
 }
 
 temp<-subset(df1,df1[,"Source.Descrip"]==controls)

 plot<-ggplot(temp, aes(x = Source.Descrip, y = PercDiff)) +
     geom_bar(stat = "identity", position = "dodge") +
     labs(title=paste("%Diff: ",genus,"- ",dbs,"- ",controltype,sep=""),
          x="Wall Type",
          y="Percent") +
     facet_wrap(~ Taxa) +
     theme_bw() +
     theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 45, hjust = 0.5, vjust = 0.5),
           axis.text.y = element_text(colour = "grey20", size = 12),
           text = element_text(size = 16))
 
 tiff(paste(output_location_seq,"\\Graphs\\bar_",genus,"_",dbs,"_",controltype,".tiff",sep=""),compression="lzw")
 grid.draw(plot)
 dev.off()
}

#Family
#GreenGenes to Silva119
setdiff(taxa_Family_greengenes,taxa_Family_silva)
temp1<-subset(taxa_Family_greengenes,select=-c(Clostridiaceae, Pseudomonadaceae))
temp2<-subset(taxa_Family_silva,select=-c(Listeriaceae, Clostridiaceae1))

#GreenGenes to Silva132
setdiff(taxa_Family_greengenes,taxa_Family_silva132)
temp3<-subset(taxa_Family_greengenes,select=-c(Clostridiaceae, Verrucomicrobiaceae))
temp4<-subset(taxa_Family_silva132,select=-c(Akkermansiaceae, Clostridiaceae1))

#Silva119 to Silva132
setdiff(taxa_Family_silva,taxa_Family_silva132)
temp5<-subset(taxa_Family_silva,select=-c(Listeriaceae, Verrucomicrobiaceae))
temp6<-subset(taxa_Family_silva132,select=-c(Akkermansiaceae, Pseudomonadaceae))

#Genus
#GreenGenes to Silva119
setdiff(taxa_Genus_greengenes,taxa_Genus_silva)
temp7<-subset(taxa_Genus_greengenes,select=-c(Rhodobacter, Roseburia, Clostridium, Escherichia, Unknown, Pseudomonas))
temp8<-subset(taxa_Genus_silva,select=-c(uncultured, Clostridiumsensustricto1, Alistipes, Listeria, Pseudobutyrivibrio, Subdoligranulum, IncertaeSedis))
temp8<-temp8[,-12] #remove Escherichia-Shigella

#GreenGenes to Silva132
setdiff(taxa_Genus_greengenes,taxa_Genus_silva132)
temp9<-subset(taxa_Genus_greengenes,select=-c(Propionibacterium, Ruminococcus, Clostridium, Unknown, Prevotella, Coprococcus,Clostridium,Escherichia ))
temp10<-subset(taxa_Genus_silva132,select=-c(Cutibacterium, Prevotella9, Alistipes, Subdoligranulum, Coprococcus2, Ruminococcus1, RuminococcaceaeNK4A214group, Prevotella7, Ruminococcus2, Ruminiclostridium5,Clostridiumsensustricto1))
temp10<-temp10[,-c(1,12,24:26)] #remove Escherichia-Shigella, [Eubacterium]halliigroup,RuminococcaceaeUCG-005, Clostridiumsensustricto1, RuminococcaceaeUCG-002, RuminococcaceaeUCG-013.
setdiff(temp9,temp10)

#Silva119 to Silva132
setdiff(taxa_Genus_silva,taxa_Genus_silva132)
temp11<-subset(taxa_Genus_silva,select=-c(uncultured, Prevotella, Listeria, Pseudobutyrivibrio, Ruminococcus, Propionibacterium, IncertaeSedis, Coprococcus))
temp12<-subset(taxa_Genus_silva132,select=-c(Prevotella9, Rhodobacter, Roseburia, Ruminiclostridium5, Cutibacterium, Coprococcus2, Ruminococcus1, RuminococcaceaeNK4A214group, Pseudomonas, Prevotella7, Ruminococcus2))
temp12<-temp12[,-c(1,23,24,25)] #Remove  RuminococcaceaeUCG-013,RuminococcaceaeUCG-002,RuminococcaceaeUCG-005, [Eubacterium]halliigroup

#Run bar plots for each comparison
database_compare<-c("GG-Silva119","GG-Silva132","Silva119-Silva132")
control_type<-c("MSA","Zymo")
count=1

for (a in taxa_levels){
 taxa_db_diffs<-taxa_diffs(get(paste("temp",count,sep="")),get(paste("temp",count+1,sep="")))
 
 taxa_db_diffs<-subset(taxa_db_diffs,Taxa!="HigherGenus" & Taxa!="HigherFamily") #View the data without the "higher categories"
 
 for (b in database_compare){
  for (c in control_type){
   bar_percdiff(taxa_db_diffs,a,b,c)
  }
  count=count+2
 }
}

#Remove all temps
remove(temp1,temp2,temp3,temp4,temp5,temp6,temp7,temp8,temp9,temp10,temp11,temp12)
```


################## Adapted from 2. Baseline_Seq
#Taxonomy Reference
```{r}
#Read in controls taxonomy file, grouped by taxa
count=1
for (a in taxa_levels){
 expected_name<-paste("expected_taxa_",a,sep="")

 assign(expected_name,read.table(paste(control_location,"Taxonomy_",a,"_2019.txt",sep=""),sep="\t",header=TRUE,row.names=count+6))
 
 write.csv(get(expected_name),paste(output_location_seq,"\\Taxa\\Taxonomy",a,"_reference_2019.csv",sep=""))
 
 assign(expected_name,get(expected_name)[-1])
 count=count+1
}
```

#Observed to Expected Counts
```{r}
tp_fp_fn_samplelevel<- function(db,tl,ct){
 sample_list<-intersect(rownames(get(paste("taxa",tl,db,sep="_"))),rownames(subset(metadata,Source.Descrip==ct)))
 exp_taxa_list<-rownames(get(paste("expected_taxa",tl,sep="_"))[ get(paste("expected_taxa",tl,sep="_"))[[ct]] !='A' , ])
 unexp_taxa_list<-setdiff(colnames(get(paste("taxa",tl,db,sep="_"))),exp_taxa_list)

 for (a in sample_list){
  fp=0
  fn=0
  tp=0
  
  taxa_db<-get(paste("taxa",tl,db,sep="_"))
  
  #Run through expected taxa list
  for (b in exp_taxa_list){
   otu_count<-taxa_db[a,b]

   
   if(is.null(otu_count) || otu_count==0 || is.na(otu_count)){
    fn=fn+1

   } else if (otu_count>0){
    tp=tp+1
   }
  }
  
  #Run through all taxa not expected
  for (b in unexp_taxa_list){
   otu_count<-taxa_db[a,b]
   
   if(otu_count==0){
    next;
   } else if (otu_count>0){
    fp=fp+1
   }
  }
  
  temp[a,"ControlType"]<-ct
  temp[a,paste("TP",db,tl,sep="_")]<-tp
  temp[a,paste("FP",db,tl,sep="_")]<-fp
  temp[a,paste("FN",db,tl,sep="_")]<-fn
  temp[a,paste("TAR",db,tl,sep="_")]<-tp/(tp+fp)
  temp[a,paste("TDR",db,tl,sep="_")]<-tp/(tp+fn)
 }
 
 return(temp)
}

temp<-data.frame()

control_type<-c("MSA1000","MSA1001","MSA1002","MSA1003","D6305","D6306","D6311")

for(a in reference_db){
 for (b in taxa_levels){
  for (c in control_type){
   temp<-tp_fp_fn_samplelevel(a,b,c)
  }
 }
}

write.csv(temp,paste(output_location_seq,"\\Taxa\\Taxonomy_tar_tdr.csv",sep=""))

remove(temp)

```